---
title: "exp2_diag_sim_new"
format: docx
lang: en-GB
editor: 
  markdown:
      wrap: 62
---


# Sampling from experiment 1

First we can sample effects from experiment 1.

```{r}
#| include: false

library(tidyverse)
library(lsr)
library(pbapply)
library(afex)
library(gt)
library(gganimate)
library(gifski)
library(av)

save_animations <- FALSE
run_simulations <- FALSE
load_big_data <- FALSE

```

```{r load data}
#| include: false
load("../data/processed/exp1_data.rdata")
load("../data/simulations/diag_simu_sum.rdata")
if(load_big_data){ load("../data/simulations/diag_simu_raw.rdata") }
```

```{r}
#| echo: false
d <-
  exp1_d |>
  dplyr::select(id, trial_info, inducer_run, diagnostic_run, correct_response, rt, con) |>
  filter(trial_info == "Diagnostic trial" | trial_info == "Inducer trial") |>
  filter(!is.na(correct_response)) |>
  mutate(rt = as.integer(rt)) 
```



```{r loss est}
#| echo: false
d -> d_ex
loss <- list()
loss$data_trials <- nrow(d_ex)

# accur
d_ex |>
  filter( trial_info == "Diagnostic trial" | trial_info == "Inducer trial" ) |>
  group_by( id ) |>
  mutate( correct_response = ifelse( trial_info=="Inducer trial" & is.na(rt), 0, correct_response) ) |>
    #' !  Non-responses count as a wrong response & are subject to the exclusion criteria.  !
  summarise( acc = sum(correct_response, na.rm = TRUE) / length( !is.na(correct_response) ) ) |>
  filter( acc < .7 ) |>
  pull( id ) -> loss$exclude_par

loss[["exclude_par_trials"]] <- length( d_ex$rt[d_ex$id == loss$exclude_par] )
loss[["exclude_par_pct"]]    <- length( d_ex$rt[d_ex$id == loss$exclude_par] ) / loss$data_trials * 100

d_ex |> filter( !(id %in% loss$exclude_par) ) -> d_ex
# rt dev+miss
d_ex |>
  mutate(rt = as.integer(rt)) |>
  ungroup() |>
  group_by(id) |>
  mutate(
    rt_crit = ifelse( 
      trial_info == "Diagnostic trial",
      mean( rt ) + sd( rt ) * 2.5,
      NA ),
      retain_trials = ifelse(
        (rt > rt_crit & trial_info == "Diagnostic trial") | # **OR**
          (is.na(rt) & trial_info == "Diagnostic trial"),
        0, 1 ) 
    ) -> d_ex

sum(d_ex$retain_trials == 0) -> loss[["rt_sd_trials"]]
sum(d_ex$retain_trials == 0) / loss$data_trials * 100 -> loss[["rt_sd_pct"]]

d_ex |> 
  filter( retain_trials == 1 ) -> d_ex
# inducers:
d_ex |>
  mutate( valid_trials = case_when( trial_info=="Inducer trial" & correct_response==1 ~ 1,
                                    trial_info=="Inducer trial" & correct_response==0 ~ 0,
                                    trial_info=="Inducer trial" & is.na(correct_response) ~ 0,
                                      # non-responses count as a wrong response
                                    T ~ NA ) ) |>
  fill(valid_trials, .direction = "up") -> d_ex

sum(d_ex$valid_trials==0) -> loss[["inducer_fail_trials"]]
sum(d_ex$valid_trials==0) / loss$data_trials * 100 -> loss[["inducer_fail_pct"]]

d_ex |> 
  filter( valid_trials == 1 ) |> 
  ungroup() -> d_ex
```

The above table indicate the number of diagnostic trials that are left after all excluding trials. 

On average, this leaves us with:
```{r}
#| echo: false
d_ex |>
  filter(trial_info=="Diagnostic trial") |>
  group_by(id) |>
  count() |>
  ungroup() |>
  summarise(avg = mean(n)) |> gt()
```
trials.

This is a loss of approximately: `r 180/240`. Whatever post-cue run we decide, we need to estimate an expected loss of about .75. 



# Generating data

Calculate the standardized means and standard deviations, and effect sizes from experiment 1. 

```{r}
#| echo: false

d_ex |> 
  ungroup() |>
  filter(!is.na(con)) |>
  summarise(
    rt = mean(rt), 
    pe = 1 - mean(correct_response),
    .by = c(id, con)
  ) |> 
  mutate(
    rt = scale(rt),
    pe = scale(pe)
  ) |>
  pivot_wider(names_from = con, values_from = c(rt,pe)) |>
  summarise(
    rt_d_m = mean(rt_FALSE-rt_TRUE),
    rt_d_sd = sd(rt_FALSE-rt_TRUE),
    rt_d = cohensD(as.double(rt_FALSE), as.double(rt_TRUE), method="paired"),
    pe_d_m = mean(pe_FALSE-pe_TRUE),
    pe_d_sd = sd(pe_FALSE-pe_TRUE),
    pe_d = cohensD(as.double(pe_FALSE), as.double(pe_TRUE), method="paired"),
  ) -> exp1_outcome
exp1_outcome
```

only output is displayed).
